Index measures for oak decline severity using phenotypic descriptors

### 2020-02-07 manuscript/manuscript.Rmd
further changes from John comments

The interactions of key raw descriptors identified from the overall importance scores and their associations with the decline indexes was investigated by using simulated descriptor values to construct index model ($M_{PDI}$,$M_{DAI}$) response surfaces for specific example cases.
The same example cases were used as described for the LIME analyses above.
This approach was similar to response surface methodology without the need for probabilistic approximation due to the low computation time required for index calculation [@abbasi2012].

@@

A scatter plot of the calculated PDI and DAI scores is shown in Figure 1 for all `r nrow(site_corrected_analysis_suitable_data)` trees for which phenotypic descriptor data was collected.
It can clearly be seen that the healthy trees group towards the lower end of the PDI scale and the symptomatic COD and AOD groups towards the higher end.
This gave a significant difference when PDI of the control and symptomatic (AOD and COD) decline statuses were compared using an independent two sample t-test (`r glue("t({PDI_status_ttest$parameter}) = {signif(PDI_status_ttest$statistic,digits = 3) %>% abs()}, p < 0.001")`) with group means of `r signif(PDI_status_ttest$estimate1,digits = 3)` and `r signif(PDI_status_ttest$estimate2,digits = 3)` for non-symptomatic and symptomatic statuses respectively.

@@

The complex descriptor interactions within the index models ($M_{PDI}$,$M_{DAI}$) investigated by simulating phenotypic data and generating the index values from the supervised index random forest regression models. 
The index values generated by the models from these simulated phenotypic data were then visualised as response surface plots, from which descriptor interactions within the indexes can be interpreted. 
Figures 3 and 4 show the simulated index value response surfaces under a variety of different circumstances for key phenotypic descriptors that contribute highly to the PDI and DAI.

@@

An unsupervised approach was favoured in this context due to the concession that there know difficulties and inaccuracy of manual decline status classification in the field [@ferretti1997].
This is highlighted in Figure 1 where a numerous trees would be considered to have been misclassified as they fall into regions of the decline indexes associated with a different decline status (e.g. COD trees with DAI values > 0.5).
The unsupervised approach has worked in this context in aptly describing decline severity as the phenotypic descriptors selected were highly specific and well suited to describing the decline severity problem.
In situations were descriptors may well be less suited, perhaps as a result of technical limitations, the use of a semi-supervised approach may be favourable.
The unsupervised approach to derive the decline indexes would be applied to a subset of the observations on which the descriptors are known to perform well and that would be considered benchmark observations by the investigators.
Indexes could then be attained for the rest of the observations by prediction based on supervised models of the benchmark subset [@chapelle2010].
This approach would be more suitable than a fully supervised approach as it would limit the inaccuracies that would be introduced by manual classification and still allow the data driven flexibility of the unsupervised approach.

>>>>

The interactions of key raw descriptors identified with high mean decrease in accuracy importance scores and their associations with the decline indexes was investigated by using simulated descriptor values to construct index model ($M_{PDI}$,$M_{DAI}$) response surfaces for specific example cases.
The same example cases were used as described for the LIME analyses above.
This approach was similar to response surface methodology without the need for probabilistic approximation due to the low computation time required for index calculation [@abbasi2012].

@@

A scatter plot of the calculated PDI and DAI scores (see Supplementary information ##) is shown in Figure 1 for all `r nrow(site_corrected_analysis_suitable_data)` trees for which phenotypic descriptor data was collected.
It can clearly be seen that the healthy trees group towards the lower end of the PDI scale and the symptomatic COD and AOD groups towards the higher end.
This gave a significant difference when PDI of the control and symptomatic (AOD and COD) decline statuses were compared using an independent two sample t-test (`r glue("t({PDI_status_ttest$parameter}) = {signif(PDI_status_ttest$statistic,digits = 3) %>% abs()}, p < 0.001")`) with group means of `r signif(PDI_status_ttest$estimate1,digits = 3)` and `r signif(PDI_status_ttest$estimate2,digits = 3)` for non-symptomatic and symptomatic statuses respectively.

@@

The complex descriptor interactions within the index models ($M_{PDI}$,$M_{DAI}$) investigated by simulating phenotypic data and generating the index values from the supervised index random forest regression models. 
The index values generated by the models from these simulated phenotypic data were then visualised as model response surface plots, from which descriptor interactions within the indexes can be interpreted. 
Figures 3 and 4 show the simulated index value response surfaces under a variety of different circumstances for key phenotypic descriptors that contribute highly to the PDI and DAI.

@@

An unsupervised approach was favoured in this context due to the concession that there known difficulties with and inaccuracy in manual decline status classification in the field [@ferretti1997].
This is highlighted in Figure 1 where a numerous trees would be considered to have been misclassified as they fall into regions of the decline indexes associated with a different decline status (e.g. COD trees with DAI values > 0.5).
The unsupervised approach has worked in this context in aptly describing decline severity as the phenotypic descriptors selected were highly specific and well suited to describing the decline severity problem.
In situations were descriptors may well be less suited, perhaps as a result of technical limitations, the use of a semi-supervised approach may be favourable.
The unsupervised approach to derive the decline indexes would be applied to a subset of the observations on which the descriptors are known to perform well and that would be considered benchmark observations by the investigators.
Indexes could then be attained for the rest of the observations by prediction based on supervised models of the benchmark subset [@chapelle2010].
This approach would be more suitable than a fully supervised approach as it would limit the inaccuracies that would be introduced by manual classification and still allow the data driven flexibility of the unsupervised approach.
